---
title: "Benchmarking the logistic growth model"
author: "Jessica Guo"
date: "12/22/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Overview

The Bayesian version of the logistic growth model will return maximum height, minimum height, and maximum growth rate for a single cultivar. The first argument is a dataframe containing canopy height (canopy_height), growing degree days (gdd), and block if applicable (site). 

There are two versions of this function: type = "simple" will treat all blocks the same, while type = "RE" will fit a random coefficients model for each block. If there is only one block per cultivar, type = "simple" will be run automatically. 

The final argument, outdir, defines the path where files should be saved. A folder for each cultivar will be created, in which the following will be saved:
1.) Gelman diagnostics as gelman_cultivar.Rdata
2.) Posterior chains as mcmc_cultivar.Rdata
3.) Traceplots ('fuzzy catepillars') of the 3 parameters (Ymax, Ymin, and Ghalf) and the population standard deviation(sig) as trace_cultivar.jpg
4.) Summary of posterior chains as mcmc_sum_cultivar.Rdata
5.) Plot of predicted vs observed canopy heights as fit_cultivar.jpg
6.) Summary of important posteriors, model convergence, and model fit parameters as out.Rdata. This dataframe is also returned by the function. 

```{r}
library(dplyr)
source("fit_logistic_growth.R")
# devtools::install_github("eddelbuettel/rbenchmark")
library(rbenchmark)
```

Read in data. 

```{r}
set.seed(5)
season6 <- na.omit(read.table(file = "~/phenophasebbn/season6_combined.txt", sep = "\t",
                              header = TRUE,
                              stringsAsFactors = FALSE))
```

## Diagnostics: single cultivar

Randomly select a single cultivar and test both the "simple" and the "RE" models. 

```{r}
#randomly sample 1 cultivar from the season for testing model
s6_cultivars <- sample(unique(season6$cultivar), size = 1)

#subset season6 dataframe by 10 randomly selected cultivars
s6_subset <- season6 %>%  filter(cultivar %in% s6_cultivars) %>% 
  select(sitename, gdd, canopy_height, cultivar, date) %>% 
  arrange(date) %>%
  mutate(site = as.numeric(as.factor(sitename)))

benchmark(
  "simple" = {
  out <- fit_logistic_growth(data = s6_subset,
                           type = "simple",
                           outdir = "~/phenophasebbn/jags/tests")
  },
  "RE" = {
  out2 <- fit_logistic_growth(data = s6_subset,
                           type = "RE",
                           outdir = "~/phenophasebbn/jags/tests")
  },
  replications = 1,
  columns = c("test", "replications", "elapsed",
            "relative", "user.self", "sys.self")
)

```

## Diagnostics: 10 cultivars

```{r}

s6_cultivars <- sample(unique(season6$cultivar), size = 10)

# Loop to subset dataframe by cultivar and run the RE model
out10 <- c()
st <- proc.time()
for(i in 1:10){
  sub <- season6 %>%  filter(cultivar %in% s6_cultivars[i]) %>% 
  select(sitename, gdd, canopy_height, cultivar, date) %>% 
  arrange(date) %>%
  mutate(site = as.numeric(as.factor(sitename)))
  
  temp <- fit_logistic_growth(data = sub,
                           type = "RE",
                           outdir = "~/phenophasebbn/jags/tests")
  out10 <- rbind(out10, temp)
}
en <- proc.time()
dur <- (en - st) /60
# user = 4.9 minutes; elapsed = 5.2 minutes
```


## Diagnostics: 10 cultivars

```{r}

s6_cultivars <- sample(unique(season6$cultivar), size = 10)

# Loop to subset dataframe by cultivar and run the RE model
out10 <- c()
st <- proc.time()
for(i in 1:10){
  sub <- season6 %>%  filter(cultivar %in% s6_cultivars[i]) %>% 
  select(sitename, gdd, canopy_height, cultivar, date) %>% 
  arrange(date) %>%
  mutate(site = as.numeric(as.factor(sitename)))
  
  temp <- fit_logistic_growth(data = sub,
                           type = "RE",
                           outdir = "~/phenophasebbn/jags/tests")
  out10 <- rbind(out10, temp)
}
en <- proc.time()
dur <- (en - st) /60
# user = 4.9 minutes; elapsed = 5.2 minutes
