---
title: "Running logistic growth by cultivar"
author: "Jessica Guo"
date: "1/4/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Overview
This script applies a hierarchical Bayesian model for logistic growth to obtain 3 parameters for each cultivar: maximum and minimum height (cm), and maximum growth rate (cm/gdd). 


Load libraries
```{r}
library(dplyr)
source("fit_logistic_growth.R")
library(ggplot2)
library(forcats)
```

Read in MAC seasons 4 and 6 cleaned height data. 
```{r}
season4 <- na.omit(read.table(file = "~/phenophasebbn/season4_combined.txt", sep = "\t",
                              header = TRUE,
                              stringsAsFactors = FALSE))

summary(c(table(season4$cultivar)))
season6 <- na.omit(read.table(file = "~/phenophasebbn/season6_combined.txt", sep = "\t",
                              header = TRUE,
                              stringsAsFactors = FALSE))
summary(c(table(season6$cultivar)))
```

## Running models with lapply for all cultivars
```{r}
# Function to turn site into numeric within cultivar
add_site <- function(x){
  x$site <- as.numeric(as.factor(x$sitename))
  return(x)
}

# split into list of 217 cultivars and add sites
s4 <- subset(season4, select = c(sitename, gdd, canopy_height, 
                                 cultivar, date))[order(season4$date),] 
s4.list <- split(s4, f = s4$cultivar)
s4.list <- lapply(s4.list, FUN = add_site)

s6 <- subset(season6, select = c(sitename, gdd, canopy_height, 
                                 cultivar, date))[order(season6$date),] 
s6.list <- split(s6, f = s6$cultivar)
s6.list <- lapply(s6.list, FUN = add_site)

# run for season 4
start <- proc.time()
out.s4 <- lapply(s4.list, FUN = fit_logistic_growth, 
                        type = "RE", outdir = "~/phenophasebbn/jags/season4")
end <- proc.time()
dur <- (end-start)[[3]]/60/60

# run for season 6
start <- proc.time()
out.s6 <- lapply(s6.list, FUN = fit_logistic_growth, 
                        type = "RE", outdir = "~/phenophasebbn/jags/season6")
end <- proc.time()
dur <- (end-start)[[3]]/60/60
```

### Combine and plot
```{r}
# from saved model outputs
cultivars <- list.files("season4")
out.s4 <- list()
for (i in 1:length(cultivars)){
  load(file.path("season4", cultivars[i], "out.Rdata"))
  out <- out %>%
    mutate(season = "season_4")
  out.s4[[i]] <- out
}
cultivars <- list.files("season6")
out.s6 <- list()
for (i in 1:length(cultivars)){
  load(file.path("season6", cultivars[i], "out.Rdata"))
  out <- out %>%
    mutate(season = "season_6")
  out.s6[[i]] <- out
}

# combine into single dataframe
s4.df <- do.call(rbind, out.s4)
s6.df <- do.call(rbind, out.s6)

all.df <- rbind(s4.df, s6.df)

#output
out.csv <- all.df %>%
  select(season, cultivar, type,
         Ymax.median, # Ymax.lower, Ymax.upper,
         Ymin.median, # Ymin.lower, Ymin.upper,
         Ghalf.median, # Ghalf.lower, Ghalf.upper,
         r2) %>%
  rename(genotype = cultivar, method_type = type,
         max_height_cm = Ymax.median,
         min_height_cm = Ymin.median,
         max_growth_cm_gdd = Ghalf.median) %>%
  mutate(method = "Bayesian logistic model of canopy height vs. gdd") %>%
  relocate(method_type, .after = last_col())

write.csv(out.csv, file = "mac_growth_rate_modeled.csv", row.names = F)

all.df %>%
  mutate(cultivar = fct_reorder(cultivar, Ymax.median)) %>%
  ggplot(aes(x = cultivar, y = Ymax.median)) + 
  geom_point(aes(color = season)) +
  # geom_errorbar(aes(ymin = Ymax.lower, ymax = Ymax.upper), alpha = 0.25) +
  scale_y_continuous("Ymax (cm)") +
  theme_bw(base_size = 12) +
  coord_flip()

all.df %>%
  mutate(cultivar = fct_reorder(cultivar, Ymin.median)) %>%
  ggplot(aes(x = cultivar, y = Ymin.median)) + 
  geom_point(aes(color = season)) +
  # geom_errorbar(aes(ymin = Ymin.lower, ymax = Ymin.upper), alpha = 0.25) +
  scale_y_continuous("Ymin (cm)") +
  theme_bw(base_size = 12) +
  coord_flip()

all.df %>%
  mutate(cultivar = fct_reorder(cultivar, Ghalf.median)) %>%
  ggplot(aes(x = cultivar, y = Ghalf.median)) + 
  geom_point(aes(color = season)) +
  # geom_errorbar(aes(ymin = Ghalf.lower, ymax = Ghalf.upper), alpha = 0.25) +
  scale_y_continuous("Ghalf (cm/gdd)") +
  theme_bw(base_size = 12) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  coord_flip()
```

KSU data is much sparser than MAC. Will the model work? Plot and test. 
```{r}
ksu <- na.omit(read.table(file = "~/phenophasebbn/ksu_combined.txt", sep = "\t",
                              header = TRUE,
                              stringsAsFactors = FALSE))
summary(c(table(ksu$cultivar)))

set.seed(8675309)
cultivar <- sample(unique(ksu$cultivar), size = 1)
#subset season6 dataframe by 10 randomly selected cultivars
ksu_subset <- ksu %>%  filter(cultivar %in% "PI570373") %>% 
  select(sitename, gdd, canopy_height, cultivar, date) %>% 
  arrange(date) %>%
  mutate(site = as.numeric(as.factor(sitename)))

ggplot(data = ksu_subset, aes(gdd, canopy_height, color = cultivar, group = sitename)) +
  geom_point() +
  geom_smooth()
```

